Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-07-15 response_to_referee.tex

\begin{document}

\maketitle

We appreciate your insightful comments.  In the following, we recap each comment followed by our response summary. In addition, we have added a few changes in the manuscript.

\begin{enumerate}
    \item \textbf{Comment:} \textit{The example on page 6 is not very clear to me. What is the meaning of the ``dscretize
beam position'' here? Where can we see that? What is $f$? I think there should be a
bit more explanations.}

@@

\textbf{Our response:} We agree that the example on page 6 could have additional clarification. In the revised manuscript, we have added an image in Fig.~2 and  substantially more details on the problem.

    \item \textbf{Comment:} \textit{The paper is essentially all about estimating the expectation of a function by QMC sampling.
But (R)QMC is also used for function approximation, simulating Markov chains,
density estimation, etc. This would required extra (connected) software tools. For example
to maintain and sort and array of Markov chains in Array-RQMC [102, 103].
A lot of this is actually implemented in the SSJ library [101], which also implements
many QMC point sets, sequences, and randomizations in the “hups” and “mcqmctools”
packages. Maybe this could be mentioned briefly somewhere.}

@@

\textbf{Our response:} In Section 8, we have added these additional problems and important tools as future work for any QMC software, including QMCPy.

    \item \textbf{Comment:} \textit{Refer.\ [9]: Maybe the authors can give the reference to the MCQMC 2020 paper on
LatNet Builder instead, because it better explains what the software is doing.}

@@

       \textbf{Our response:} We have replaced the original reference with the suggested one, which is now listed as reference [37] in the revised manuscript.

    \item \textbf{Comment:} \textit{Section 5: Although Python is very popular and easy to use, perhaps it should be
recognized explicitly that it is not the best tool if we want speed. I made the following
simple experiment to compare the speeds of qmcpy with that of SSJ. We want to
integrate the simple function $f(u_1, u_2) = u_1 + u_2 - 1$ over the unit square, using
RQMC with Sobol points $+$ a linear scrambling $+$ a random digital shift, say with
$n = 214$ points. I computed $20,000$ independent replicates of the RQMC estimator,
with independent randomizations. With SSJ, this took 4.7 seconds on my Thinkpad
laptop. With qmcpy, it took about two minutes, which is about 25 times more. I
am not asking that the authors report this experiment in their paper, but I think
they should mention that Python can be significantly slower than a good Java or
\text{C++} implementation. One possibility would be to put a Python layer over a fast implementation.
}

### 2023-07-17 response_to_referee.tex

\begin{enumerate}
    \item \textbf{Comment:} \textit{The example on page 6 is not very clear to me. What is the meaning of the ``dscretize
beam position'' here? Where can we see that? What is $f$? I think there should be a
bit more explanations.}

\textbf{Our response:} We agree that the example on page 6 could have additional clarification. In the revised manuscript, we have added an image in Fig.~2 and  substantially more details on the problem.

    \item \textbf{Comment:} \textit{The paper is essentially all about estimating the expectation of a function by QMC sampling.
But (R)QMC is also used for function approximation, simulating Markov chains,
density estimation, etc. This would required extra (connected) software tools. For example
to maintain and sort and array of Markov chains in Array-RQMC [102, 103].
A lot of this is actually implemented in the SSJ library [101], which also implements
many QMC point sets, sequences, and randomizations in the “hups” and “mcqmctools”
packages. Maybe this could be mentioned briefly somewhere.}

\textbf{Our response:} In Section 8, we have added these additional problems, summarized some results so far, and proposed future work in this direction for  QMC software, including QMCPy.

### 2023-07-17 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: Signed off
Updates from Overleaf

%%%%%%% Author sign off on this revised submission, last revision 7/12/2023
% Sou-Cheng Choi signed off on 
% Yuhan Ding signed off on 
% Fred J. Hickernell signed off on 
% Jagadeeswaran Rathinavel signed off on 7/13/2023
% Aleksei Sorokin signed off on 7/12/2023

@@

A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
\begin{enumerate}
    \item Draw a sequence of $n$ low-discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
    \item Estimate the true mean, $\mu$, by the sample mean,
    \begin{equation}
        \hat{\mu}_n := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
        \label{eq:mc_average}
    \end{equation}
\end{enumerate}
However, the practice of QMC is often more complicated.  The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with a small number of sampling points, $n$.  Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.

@@

\section{Sustainable} \label{CDHJS_sec:sustainable}
Great QMC software libraries should be alive.  They should be under active development or merged into other libraries that are actively maintained.  Commercial software has an infrastructure to maintain their algorithms, but these algorithms need periodic refreshing. Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.

>>>>

%%%%%%% Author sign off on this revised submission, last revision 7/12/2023
% Sou-Cheng Choi signed off on 7/15/2023
% Yuhan Ding signed off on 
% Fred J. Hickernell signed off on 
% Jagadeeswaran Rathinavel signed off on 7/13/2023
% Aleksei Sorokin signed off on 7/12/2023

@@

A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
\begin{enumerate}
    \item Draw a sequence of $n$ low discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
    \item Estimate the true mean, $\mu$, by the sample mean,
    \begin{equation}
        \hat{\mu}_n := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
        \label{eq:mc_average}
    \end{equation}
\end{enumerate}

However, the practice of QMC is often more complicated.  The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with a small number of sampling points, $n$.  Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.

@@

\section{Sustainable} \label{CDHJS_sec:sustainable}
Great QMC software libraries should be awake, not dormant.  They should be under active development or merged into other libraries that are actively maintained.  Commercial software has an infrastructure to maintain their algorithms, but these algorithms need periodic refreshing. Open source software requires a community of active users and developers to keep the library up-to-date and bug-free.
